---
title: "DATA605_HW11_Simple Regression Analysis"
author: "Charles Ugiagbe"
date: "4/10/2022"
output:
  rmdformats::readthedown:
    gallery: no
    highlight: tango
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```


## Question

Using the “cars” dataset in R, build a linear model for stopping distance as a function of speed and
replicate the analysis of your textbook chapter 3 (visualization, quality evaluation of the model, and
residual analysis.)

###    Solution

```{r}
glimpse(cars)
```

Dataset cars includes 50 observations with 2 variables - speed and dist(distance) as shown in the glimpse above


##   Visualization

```{r}
plot(cars$speed,cars$dist, main="Distance vs Speed",
xlab="Speed", ylab="Dist")
```

##   The Linear Model Function

The simplest regression model is a straightline. It has the mathematical form:

$ŷ = b_{0} + b_{1}x$

where:

$x$ is the input to the system,

$b_{0}$ is the y-intercept of the line, 

$b_{1}$ is the slope, and 

$ŷ$ is the output value the model predicts. The ^ indicates a predicted or estimated value, not the actual observed value.


```{r}
cars_lm <- lm(cars$dist ~ cars$speed)
cars_lm
```

In this case, the y-intercept is b_{0} = -17.579 and the slope is b_{1} = 3.932. Thus, the final regression model is:

$y = -17.579 + 3.932x$

```{r}
plot(cars$speed, cars$dist, xlab='Speed', ylab='Distance', main='Cars Linear Regression Model')
abline(cars_lm)
```


```{r}
summary(cars_lm)
```

##   Residual Analysis

```{r}
plot(fitted(cars_lm),resid(cars_lm), main="Residuals")
abline(0, 0)
```

```{r}
qqnorm(resid(cars_lm))
qqline(resid(cars_lm))
```

```{r}
par(mfrow=c(2,2))
plot(cars_lm)
```

**From the residual vs fitted value plot we can see the there is no definitepattern in the data hence the data randomness of the residuals and heteroscidatcity is satisfied.**

**From the normal q-q plot we can see that the residuals are Approximately normally distributed.**

**From the overall analysis we can say that the model is a well fitted model since the assumptions of the linear regression model are satisfied here.**

