---
title: "DATA605_HW13_Univariate & Multivariate Calculus"
author: "Charles Ugiagbe"
date: "4/24/2022"
output:
  rmdformats::readthedown:
    gallery: no
    highlight: tango
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##    Question 1

Use integration by substitution to solve the integral below.

$\int{4e^{-7x}dx}$

##    Solution 1

Let $u=-7x$, then $du = -7dx$.
$dx = -\frac{du}{7}$

$$
\begin{split}
\int{4e^{-7x}dx} & = \int{4e^{u}\left ( \frac{-du}{7} \right )} \\
&= -\frac{4}{7}\int{e^u du} \\
&= \frac{-4}{7}e^u+constant \\
&= -\frac{4}{7}e^{-7x}+ constant
\end{split}
$$

##    Question 2

Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of $\frac{dN}{dt} = -\frac{3150}{t^4}-220$ bacteria per cubic centimeter per day, where $t$ is the number of days since treatment began. Find a function $N(t)$ to estimate the level of contamination if the level after $1$ day was $6530$ bacteria per cubic centimeter.

##    Solution 2

$$
\frac{dN}{dt} = N'(t) = \frac{-3150}{t^4}-220 \\
N(t) = \int{(\frac{-3150}{t^4}-220) dt} \\ 
= \frac{1050}{t^3}-220t+C \\
Since N(1)= 6530, then \\

\begin{split}
N(t) &= \frac{1050}{t^3}-220t+C \\
N(1) &= 6530 \\
\frac{1050}{1^3}-220\times 1 +C &= 6530 \\
C &= 6530 - 1050 + 220 \\
C &= 5700
\end{split}
$$

Since $N(1)= 6530$, then 

**The level of contamination can be estimated by the following function:** 

$N(t) = \frac{1050}{t^3}-220t+5700$


##    Question 3

Find the total area of the red rectangles in the figure below, where the equation of the line is
$f(x)=2x-9$.

##    Solution 3

Each square in the graph has an area of $1$. Each rectangle has a width of $1$. Counting the height of each rectangle mark in yellow gives  the areas as: $Area=1+3+5+7=16$.

or we can also solve it using R function

```{r}
#Find area in-build function
Q3 = function(x) {2*x -9}
```

```{r}
#Find the difference between areas under the curve
area <- integrate(Q3, 4.5, 8.5)$value
area <- round(as.numeric(area))
print(area)
```


##    Question 4

Find the area of the region bounded by the graphs of the given equations. 

$y_1 = x_1^2 - 2x_1-2$\
$y_2 = x_2 + 2$

##    Solution 4

To get the area of the region bounded by the two equation, we first solve the two equation to get the upper and the lower limit.

$y=x^{2}−2x−2$\  $y=x+2$

we equate and solve both as quadratic equation to get be bounded region.

$x^{2}−2x−2 = x+2$

$x^{2}−2x-x-2−2 = 0$

$x^{2}−3x-4 = 0$

$(x+1)(x-4) = 0$

$x = -1$ or $x = 4$

The upper limit of the integral is 4 while the lower limit is -1


```{r}
#Find area in-build function
f1 = function(x) {x + 2}
f2 = function(x) {x^2 -2*x -2}
```

```{r}
#Find the difference between areas under the curve
area1 <- integrate(f1, -1, 4)
area2 <- integrate(f2, -1, 4)
area <- round((area1$value - area2$value),4)
print(area)
```


##    Question 5

A beauty supply store expects to sell 110 flat irons during the next year. It costs \$3.75 to store one flat iron for one year. There is a fixed cost of $8.25 for each order. Find the lot size and the number of orders per year that will minimize inventory costs. 

##    Solution 5

Let $x$ be a number of flat irons to order. 

$Yearly\ storage\ cost = {Storage\ cost\ per\ iron} \times {Average\ number\ of\ irons\ stored} = 3.75 \times x/2 = 1.875x$

$Yearly\ ordering\ cost = {Cost\ of\ each\ order} \times {Number\ of\ orders} = 8.25 \times 110/x = 907.5/$

$Inventory\ cost = Yearly\ storage\ cost + Yearly\ ordering\ cost = 1.875x+907.5/x = f(x)$

To find the minimized value, differentiate and solve at $0$:

$$
\begin{split}
f'(x) &= 1.875-\frac{907.5}{x^2} \\
f'(x) &= 0 \\ 
1.875-\frac{907.5}{x^2} &= 0 \\
1.875&= \frac{907.5}{x^2} \\
1.875x^2&= 907.5 \\
x^2&= \frac{907.5}{1.875} \\
x&= \sqrt{\frac{907.5}{1.875}} \\
x&=\sqrt{484} \\
x&=22
\end{split}
$$

Each order should contain $22$ flat irons, so there should be $110/22=5$ orders. 

##    Question 6

Use integration by parts to solve the integral below. 

$\int{ln(9x) \times x^6 dx}$


##    Solution 6 

Let $u= ln(9x)$, then $\frac{du}{dx}=\frac{1}{x}$.

Let $\frac{dv}{dx}=x^6$, then $v = \int{x^6 dx} = \frac{1}{7}x^7$.

Using the formula for integration by parts: $\int{u \frac{dv}{dx}dx} = uv - \int{v \frac{du}{dx} dx}$

$$
\begin{split}
\int{ln(9x) \times x^6 dx} &= \frac{1}{7}x^7 \times ln(9x) - \int{\frac{1}{7}x^7 \times \frac{1}{x} dx} \\
&=\frac{1}{7}x^7 \times ln(9x) - \int{\frac{1}{7}x^6 dx} \\
&=\frac{7}{49}x^7 \times ln(9x) - \frac{1}{49}x^7 + constant \\
&=\frac{1}{49}x^7 (7ln(9x) - 1) + constant \\
\end{split}
$$


##    Question 7

Determine whether $f(x)$ is a probability density function on the interval $[1, e^6]$. If not, determine the value of the definite integral. 

$f(x) = \frac{1}{6x}$

##    Solution 7 


$$
\begin{split}
\int_1^{e^6}\frac{1}{6x} dx &= \frac{1}{6} ln(x)|_1^{e^6} \\
&= \frac{1}{6} ln(e^6) - \frac{1}{6} ln(1) \\
&= \frac{1}{6} \times 6 - \frac{1}{6} \times 0 \\
&= 1
\end{split}
$$
**The definite integral of the function on interval $[1, e^6]$ is $1$. Additionally, if $x>0$, then $f(x)>0$, so for this interval $f(x)>0$. As long as $f(x)=0$ outside of the given interval, this satisfies PDF requirements and this function is a probability density function**. 