---
title: "DATA605_HW3_Eigenvalues and Eigenvectors"
author: "Charles Ugiagbe"
date: "2/11/2022"
output:
  rmdformats::readthedown:
    self_contained: yes
    thumbnails: yes
    lightbox: yes
    gallery: no
    highlight: tango
    code_folding: show
editor_options: 
  chunk_output_type: inline
---

```{r message=FALSE, warning=FALSE}
library(pracma)
```



#   Problem set 1

(1) What is the rank of the matrix A?

$$
A = \begin{bmatrix}
1 & 2 & 3 & 4 \\
-2 & 0 & 1 & 3 \\
0 & 1 & -2 & 1 \\
5 & 4 & -2 & -3 \\
\end{bmatrix}
$$
 

##    Solution 1

```{r}
A <- matrix(c(1,-2,0,5,2,0,1,4,3,1,-2,-2,4,3,1,-3), ncol=4)
A
```

```{r}
det(A)
```

Since $Det(A)\neq 0$, we say there is no linear dependence of any role or column.

Rank of a matrix A is the number of role or column that are not a linear combination of another role of column. 

```{r}
qr(A)$rank
```


(2) Given an mxn matrix where m > n, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?

##    Solution 2

For a Non-Zero Matrix, the Maximum rank can't be greater than the the smaller dimension of the matrix which is n and the mininum rank can't be less than 1.  
Therefore, the rank is between 1 and n (inclusive).
Maximum Rank = n
Mininum Rank = 1



(3) What is the rank of matrix B? 

$$ B = \begin{bmatrix}
1 & 2 & 1 \\
3 & 6 & 3 \\
2 & 4 & 2 \\
\end{bmatrix}$$


##   Solution 3

```{r}
B <- matrix(c(1,3,2,2,6,4,1,3,2), nrow = 3)
B
```


```{r}
det(B)
```

The rank of B is 1, because:

(i) The determinant of this square matrix is 0, which hints at linear dependence.
(ii) Each row and column is a linear combination of other rows and column. We manipulate the matrix to get the number of NOn-zero role to be 1. Hence, the rank is 1



#   Problem Set 2

Compute the eigenvalues and eigenvectors of the matrix A. You’ll need to show your work. You’ll need to write out the characteristic polynomial and show your solution.

$A = \begin{bmatrix}
1 & 2 & 3 \\
0 & 4 & 5 \\
0 & 0 & 6 \\
\end{bmatrix}$

##  Solution
                                                       
$Det (A - \lambda I)= 0$

$Det\left ( \begin{bmatrix}
1 & 2 & 3 \\
0 & 4 & 5 \\
0 & 0 & 6 \\
\end{bmatrix}- \lambda \begin{bmatrix}
1 & 0 & 0 \\
0 & 1 & 0 \\
0 & 0 & 1 \\
\end{bmatrix} \right ) = 0$

$= Det\begin{bmatrix}
1-\lambda  & 2 & 3 \\
0 & 4-\lambda  & 5 \\
0 & 0 & 6-\lambda  \\
\end{bmatrix}$


$= (1-\lambda )((4-\lambda)(6-\lambda)-0\times 5) - 2((6-\lambda)\times 0-0\times 5) + 3((0\times 0)-(0\times (4-\lambda)))$

simplifying the above equation, we get the characteristic polynomial as:

$P(\lambda )= -\lambda ^3 + 11\lambda^2 - 34\lambda + 24$ 

Equating the characteristiv polynomial to zero(0) and Solving to get the Eigenvalues, we get:

$(-\lambda + 6)(\lambda - 4)(\lambda - 1) = 0$ 

Eigenvalues of matrix A are: $\lambda = 1, \lambda = 4, \lambda = 6$


To calculate Eigenvectors, we use the formular $(AX-\lambda I) = 0$

using R to reduce to matrix

```{r}
A <- matrix(c(1,2,3,0,4,5,0,0,6), nrow=3, byrow=TRUE)
```

If $\lambda=1$, then $A - 1I_3$ is row-reduced to

```{r}
rref(A - 1 * diag(3))
```

$$
 \begin{bmatrix}
 0 &1 &0\\
 0 &0 &1\\
 0&0&0
 \end{bmatrix}
 \begin{bmatrix}
 v_1\\
 v_2\\
 v_3
 \end{bmatrix}
 =
 \begin{bmatrix}
 0\\
 0\\
 0
 \end{bmatrix}
$$

Then $v_1=v_1$ and $v_2=0$ and $v_3=0$. The **eigenvector** for $\lambda=1$ is  

$$
 X_{1}=
 \begin{bmatrix}
 1\\
 0\\
 0
 \end{bmatrix}
$$

If $\lambda=4$, then $A - 4I_3$ is row-reduced to 

```{r}
rref(A - 4 * diag(3))
```

$$
 \begin{bmatrix}
 1 &-\frac{2}{3} &0\\
 0 &0 &1\\
 0&0&0
 \end{bmatrix}
 \begin{bmatrix}
 v_1\\
 v_2\\
 v_3
 \end{bmatrix}
 =
 \begin{bmatrix}
 0\\
 0\\
 0
 \end{bmatrix}
$$

Then $v_1 - \frac{2}{3}v_2=0$ and $v_3=0$. 

Or $v_1=v_1$ and $v_2=\frac{3}{2}v_1=1.5v_1$ and $v_3=0$. 

The **eigenvector** for $\lambda=4$ is  

$$
 X_{2}=
 \begin{bmatrix}
 1\\
 1.5\\
 0
 \end{bmatrix}
$$

Finally, if $\lambda=6$, then $A - 6I_3$ is row-reduced to 

```{r}
rref(A - 6 * diag(3))
```

$$
 \begin{bmatrix}
 1 &0 &-1.6\\
 0 &1 &-2.5\\
 0&0&0
 \end{bmatrix}
 \begin{bmatrix}
 v_1\\
 v_2\\
 v_3
 \end{bmatrix}
 =
 \begin{bmatrix}
 0\\
 0\\
 0
 \end{bmatrix}
$$

Then $v_1-1.6v_3=0$ and $v_2-2.5v_3=0$. 

Or $v_1=1.6v_3$ and $v_2=2.5v_3$ and $v_3=v_3$. 

The **eigenvector** for $\lambda=6$ is:  

$$
 X_{3}=
 \begin{bmatrix}
 1.6\\
 2.5\\
 1
 \end{bmatrix}
$$

Eigenvalues of matrix A are: $\lambda_{1} = 1, \lambda_{2} = 4, \lambda_{3} = 6$

Eigenvectors of matrix A are: $X_{1}=\begin{bmatrix}1\\0\\0\end{bmatrix}$ , $X_{2}=\begin{bmatrix}1\\1.5\\0\end{bmatrix}$ and $X_{3}=\begin{bmatrix}1.6\\2.5\\1\end{bmatrix}$