---
title: "DATA605_HW10_Markov Chains & Simple Ramdom Walk"
author: "Charles Ugiagbe"
date: "4/2/2022"
output:
  rmdformats::readthedown:
    self_contained: yes
    thumbnails: yes
    lightbox: yes
    gallery: no
    highlight: tango
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem Set

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars. A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6. Find the probability that he wins 8 dollars before losing all of his money if

(a)he bets 1 dollar each time (timid strategy).

(b)he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8  dollars (bold strategy).

(c)Which strategy gives Smith the better chance of getting out of jail?

##    Solution (a)

Using Gambler ruin Probability method

$P_{z} = \frac{1-(q/p)^{z}}{1 - (q/p)^{m}}$ $where, q \neq  p , 0 \leq Z \leq M$ 

and the quantity $P_{z}$ to be the probability that the gambler’s stake reaches M without ever having reached 0.

```{r}
z = 1
M = 8
p = 0.4
q = 0.6

p = (1-(q/p)^z) / (1-(q/p)^M)
p
```
**with the Timid Strategy,  the probability that he wins 8 dollars before losing all of his money if he bets 1 dollar each time is 0.0203**

## Solution (b)

If he uses the bold strategy, it means that he would bet all the money each time he wins. So, in this case there are only 4 steps 1, 2 (1st win), 4 (2nd win), 8 (3rd win) to reach $8. 

Because this is bold strategy, we can also solve this problem by Binomial distribution besides the Absorbing Markov chain.

Binomial method:

Number of trials, n = 3

Number of successes, x = 3

$_{x}^{n}\textrm{C}p^{x}(1-p)^{n-x}$

$_{3}^{3}\textrm{C}*0.4^{3}*(1-0.4)^{3-3}$

= 0.064

##    Part (c)

The bold strategy gives Smith approximately 3 times better chance of getting out of jail with the probability 0.064 than the smaller probability 0.0203 given by the timid strategy.


