---
title: "DATA605_HW7_DISTRIBUTIONS AND DENSITIES"
author: "Charles Ugiagbe"
date: "3/15/2022"
output:
  rmdformats::readthedown:
    self_contained: yes
    thumbnails: yes
    lightbox: yes
    gallery: no
    highlight: tango
editor_options: 
  chunk_output_type: inline
---

###   Question 1
Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y.

##  Solution 1

Assuming that each $X_{i}$ has k possibilities, we will have $k^{n}$ possible values in total

1 is the minimum of the Xi’s, the nymber of ways of getting Y.

$P(x = 1) = \frac{k^{n} - (k-1)^{n}}{k^{n}}$

$P(x = 2) = \frac{(k-2+1)^{n} - (k-2)^{n}}{k^{n}}$

For $X_{1}, X_{2}, . . . , X_{n}$ mutually independent random variables, If Y = j  then, the minimum value j is given as:

$P(x = j) = \frac{(k-j+1)^{n} - (k-j)^{n}}{k^{n}}$


###   Question 2
Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.).

2a.What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years..)

##   Solution 2A (Geometric)

Geometric distribution is given as:

$P_{r}(X=k) = (1−p)^{k-1}p$

```{r}
p <- 1 / 10
geometric_prob <- ((1 - p)^ (8 - 1) * p)
round(geometric_prob, 4)
```
**Expected value**
```{r}
EV_geometric <- 1 / p
EV_geometric
```
**Standard Deviation**
```{r}
sd_geometric <- sqrt((1 - p) / (p ^ 2))
round(sd_geometric, 4)
```

2b. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential.

##   Solution 2B (Exponential)

The exponential distribution is given as:

$f(x) = \lambda e^{\lambda x}$

```{r}
lambda <- 1 / 10
e <- exp(1)
exponential_prob <- lambda*e ^ (- lambda * 8)
round(exponential_prob, 4)
```

**Expected value**
```{r}
EV_exponential <- 1 / lambda
EV_exponential
```

**Standard Deviation**
```{r}
sd_exponential <- sqrt(1 / (lambda ^ 2))
sd_exponential
```


2c. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)

##   Solution 2C (Binomial)

The Binomial distribution is given as:

$f(r,n,p)=\binom{n}{r}p^{r}(1−p)^{n-r}$

```{r}
p <- 1 / 10
binomial_prob <- choose(8, 0) * (p ^ 0) * ((1 - p) ^ (8))
round(binomial_prob, 4)
```

```{r}
expected_value_binomial <- p * 8
expected_value_binomial
```


```{r}
n = 8
sd_binomial <- sqrt(n * p  * (1 - p))
round(sd_binomial, 4)
```


2d. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson.

##   Solution 2D (Poisson)

Poisson Distribution is given as:

$P(x,\lambda)= \frac{e^{-\lambda}\lambda^{x}}{x!}$

```{r}
lambda_poisson = 8 * (1 / 10)
e <- exp(1)
poisson_prob <-(e ^ (- lambda_poisson) * lambda_poisson ^ 0) / factorial(0)
round(poisson_prob, 4)
```
**Expected value**
```{r}
EV_poisson <- lambda_poisson
EV_poisson
```
**Standard Deviation**
```{r}
sd_poisson <- sqrt(lambda_poisson)
round(sd_poisson, 4) 
```

