---
title: "DATA605_HW9_Central Limit Theorem & Generating Functions"
author: "Charles Ugiagbe"
date: "3/27/2022"
output:
  rmdformats::readthedown:
    self_contained: yes
    thumbnails: yes
    lightbox: yes
    gallery: no
    highlight: tango
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Page 363 Exercise 11

The price of one share of stock in the Pilsdorff Beer Company is given by $Y_n$ on the $n$th day of the year. Finn observes that
the differences $X_n = Y_{n+1} - Y_n$ appear to be independent random variables
with a common distribution having mean $\mu = 0$ and variance $\sigma^2 = 1/4$. If
$Y_1 = 100$, estimate the probability that $Y_{365}$ is

a. $\ge 100$
b. $\ge 110$
c. $\ge 120$

### SOLUTION

Since $X_n$ is an independent random variable, then its sum $S_n$ is normally distributed. 

$$
\begin{split}
S_n &= X_1+X_2+X_3...+X_n \\
&=(Y_2-Y_1)+(Y_3-Y_2)+(Y_4-Y_3)+...+(Y_{n+1}-Y_n)\\
&=Y_2-Y_1+Y_3-Y_2+Y_4-Y_3+...+Y_{n+1}-Y_n\\
&=Y_{n+1}-Y_1\\
&=Y_{n+1}-100
\end{split}
$$

Mean of $S_n$: $\mu_{S_n} = \mu_{X_1}+\mu_{X_2}+\mu_{X_3}+...+\mu_{X_n} = n\mu_X = 0$

Variance of $S_n$: $\sigma^2_{S_n} = \sigma^2_{X_1}+\sigma^2_{X_2}+\sigma^2_{X_3}+...+\sigma^2_{X_n} = n\sigma^2_X = n \times 1/4 = n/4$

Standard deviation of $S_n$: $\sigma_{S_n} = \sqrt{n}/2$

Consider $n=364$, then $S_{364} = Y_{365} -100$ and $Y_{365} = S_{364}+100$.

$\sigma^2_{S_{364}} = 364/4 = 91$ and $\sigma_{S_{364}} = \sqrt{91}$

### Solution (a)

$$
\begin{split}
P(Y_{365} \ge 100) &= P(S_{364}+100 \ge 100) \\
&= P(S_{364} \ge 0)
\end{split}
$$

Since $S_n$ is normally distributed with mean $0$ and normal distribution is symmetric, exactly half of values will be greater than mean. 

**ANSWER:** $P(S_{364}\ge 0) =0.5$

### Solution (b)

$$
\begin{split}
P(Y_{365} \ge 110) &= P(S_{364}+100 \ge 110) \\
&= P(S_{364} \ge 10) \\
&= P(S^*_{364}\ge 10/\sqrt{91})
\end{split}
$$  

```{r}
z <- 10/sqrt(91)
pnorm(z, lower.tail=FALSE)
```

**ANSWER:** $P(Y{365} \ge 110) \approx 0.14725$

### Solution (c)

$$
\begin{split}
P(Y_{365} \ge 120) &= P(S_{364}+100 \ge 120) \\
&= P(S_{364} \ge 20) \\
&= P(S^*_{364}\ge 20/\sqrt{91})
\end{split}
$$


```{r}
z <- 20/sqrt(91)
pnorm(z, lower.tail=FALSE)
```

**ANSWER:** $P(Y{365} \ge 120) \approx 0.01802$


## Problem 2

Calculate the expected value and variance of the binomial distribution using the moment generating function.

### SOLUTION

For binomial distribution, $P(X=k) = {n \choose k} p^k q^{n-k}$

where $q=1-p$.

The moment generating function is $M_X(t)=(q+pe^t)^n$.

The first moment is $M'_X(t) = n(q+pe^t)^{n-1}pe^t$.

The expected value is the first moment evaluated at $t=0$:

$$
\begin{split}
E(X)=M'_X(0) &= n(q+pe^0)^{n-1}pe^0\\
&= n(q+p)^{n-1}p\\
&= np(1-p+p)^{n-1}\\
&= np1^{n-1}\\
&=np
\end{split}
$$

The second moment is $M''_X(t) = n(n-1)(q+pe^t)^{n-2}p^2 e^{2t}+n(q+pe^t)^{n-1}pe^t$.

Evaluate the second moment at $t=0$:

$$
\begin{split}
E(X^2)=M''_X(0) &= n(n-1)(q+pe^0)^{n-2}p^2 e^0+n(q+pe^0)^{n-1}pe^0\\
&= n(n-1)(1-p+p)^{n-2}p^2+n(1-p+p)^{n-1}p\\
&= n(n-1)p^2+np
\end{split}
$$

The variance is $V(X)=E(X^2)-E(X)^2$:

$$
\begin{split}
V(X) &= n(n-1)p^2+np-n^2p^2 \\
&= np((n-1)p+1-np) \\
&= np(np-p+1-np) \\
&= np(1-p) \\
&= npq
\end{split}
$$

We arrived at the known definitions for binomial distribution - $E(X)=np$ and $V(X)=npq$.


## Problem 3

Calculate the expected value and variance of the exponential distribution using the moment generating function.

### SOLUTION

For exponential distribution, $f(x)=\lambda e^{-\lambda x}$.

The moment generating function is $M_X(t)=\frac{\lambda}{\lambda-t}, t<\lambda$.

Using WolframAlpha, we get $M'_X(t) = \frac{\lambda}{(\lambda-t)^2}$ and $M''_X(t) = \frac{2\lambda}{(\lambda-t)^3}$.

Expected value:

$$
\begin{split}
E(X)=M'_X(0) &= \frac{\lambda}{(\lambda-0)^2} \\
&= \frac{\lambda}{\lambda^2}\\
&= \frac{1}{\lambda}
\end{split}
$$

Variance:

$$
\begin{split}
V(X) = E(X^2)-E(X)^2 &= M''_X(0)-M'_X(0)^2 \\
&=\frac{2\lambda}{(\lambda-0)^3} - \frac{1}{\lambda^2}\\
&=\frac{2\lambda}{\lambda^3} - \frac{1}{\lambda^2}\\
&=\frac{2}{\lambda^2} - \frac{1}{\lambda^2}\\
&=\frac{1}{\lambda^2}
\end{split}
$$

We arrived at the known definitions for binomial distribution - $E(X)=1/\lambda$ and $V(X)=1/\lambda^2$.
