---
title: "DATA605_Homework2"
author: "Charles Ugiagbe"
date: "2/4/2022"
output:
  rmdformats::readthedown:
    self_contained: yes
    thumbnails: yes
    lightbox: yes
    gallery: no
    highlight: tango
    code_folding: show
editor_options: 
  chunk_output_type: inline
---


```{r message=FALSE, warning=FALSE}
library(tidyverse)
```



# Problem Set 1

(1) Show that $A^{T}A \neq AA^{T}$ in general. (Proof and demonstration.)
(2) For a special type of square matrix A, we get $A^{T}A = AA^{T}$ . Under what conditions
could this be true? (Hint: The Identity matrix I is an example of such a matrix).


## Solution (1)

(1) Proof to show that $A^{T}A \neq AA^{T}$


### Proof
Suppose a matrix $A$ is an $m \times n$ matrix and a matrix $B$ is an $n \times p$ matrix. Let vectors $A_{1}, A_{2}, A_{3},...,A_{n}$ denote the columns of $A$ and let the vectors $B_{1}, B_{2}, B_{3},...,B_{p}$ denote the columns of $B$. 

Then for $1\leq i\leq m, 1\leq j\leq p$, the individual entries of $AB$ are:


$$[AB]_{ij} = [AB_{j}]_{i}$$
$$= [ [B_{j}]_{1}A_{1} + [B_{j}]_{2}A_{2} + \cdots + [B_{j}]_{n}A_{n} ]_{i}$$
$$= [ [B_{j}]_{1}A_{1}]_{i} + [[B_{j}]_{2}A_{2}]_{i} + \cdots + [[B_{j}]_{n}A_{n} ]_{i}$$
$$= [B_{j}]_{1}[A_{1}]_{i} + [B_{j}]_{2}[A_{2}]_{i} + \cdots + [B_{j}]_{n}[A_{n}]_{i}$$
$$= [B_{1j}][A]_{i1} + [B_{2j}][A]_{i2} + \cdots + [B_{nj}][A]_{in} $$
$$= [A_{i1}][B]_{1j} + [A_{i2}][B]_{2j} + \cdots + [A_{in}][B]_{nj}$$
$$= \sum_{k=1}^{n} [A]_{ik}[B]_{kj}$$

Base on the proof above, suppose the matrix $A$ is an $m \times  n$ matrix, which implies that $A^{T}$ is an $n \times m$ matrix. Then for $1\leq i\leq m, 1\leq j\leq m$, the individual entries of $AA^{T}$ are given by 
$$[AA^{T}]_{ij} = [A]_{i1}[A^{T}]_{1j} + [A]_{i2}[A^{T}]_{2j} + [A]_{i3}[A^{T}]_{3j} + \cdots + [A]_{in}[A^{T}]_{nj}$$
$$= \sum_{k=1}^{n} [A]_{ik}[A^{T}]_{kj}$$


And for $1\leq i\leq n, 1\leq j\leq n$, the individual entries of $A^{T}A$ are given by 
$$[A^{T}A]_{ij} = [A^{T}]_{i1}[A]_{1j} + [A^{T}]_{i2}[A]_{2j} + [A^{T}]_{i3}[A]_{3j} + \cdots + [A^{T}]_{im}[A]_{mj}$$
$$= \sum_{k=1}^{m} [A^{T}]_{ik}[A]_{kj}$$
, where $AA^{T}$ is an $m \times m$ matrix and $A^{T}A$ is an $n \times n$ matrix.

Therefore, $A^{T}A \neq AA^{T}$ in general.



###   Demostration.




$$
{ A }^{ T }A \neq A{ A }^{ T }
$$


Suppose we have matrix A and its transpose


Let $A=\begin{bmatrix} 2 & -1 \\ 3 & 2 \end{bmatrix}$ and 
$\\{ A }^{ T }=\begin{bmatrix} 2 & 3 \\ -1 & 2 \end{bmatrix}$

calculate $A^{T}A$ and $AA^{T}$.


Using matrix multiplication, we evaluate the matrix above

$$
{ A }^{ T }A= 
\begin{bmatrix} 2 & 3 \\ -1 & 1 \end{bmatrix}
\begin{bmatrix} 2 & -1 \\ 3 & 1 \end{bmatrix}\\
{ A }^{ T }A=
\begin{bmatrix} 4+9 & -2+3 \\ -2+3 & 1+1 \end{bmatrix}\\
{ A }^{ T }A= \begin{bmatrix} 13 & 1 \\ 1 & 2 \end{bmatrix}
$$ 

----------------------------------------------------------------------------------

$$
A{ A }^{ T }= 
\begin{bmatrix} 2 & -1 \\ 3 & 1 \end{bmatrix}
\begin{bmatrix} 2 & 3 \\ -1 & 1 \end{bmatrix}\\
A{ A }^{ T }=
\begin{bmatrix} 4+1 & 6-1 \\ 6-1 & 9+1 \end{bmatrix}\\
A{ A }^{ T }= \begin{bmatrix} 5 & 5 \\ 5 & 10 \end{bmatrix}
$$

${ A }^{ T }A= \begin{bmatrix} 13 & 1 \\ 1 & 2 \end{bmatrix}$ and $A{ A }^{ T }= \begin{bmatrix} 5 & 5 \\ 5 & 10 \end{bmatrix}$

Hence, we conclude that: ${ A }^{ T }A \neq A{ A }^{ T }$


$$
{ A }^{ T }A\neq A{ A }^{ T }
$$


## Solution (2)

For a special type of square matrix A, we get $A^{T}A = AA^{T}$ . Under what conditions
could this be true? (Hint: The Identity matrix I is an example of such a matrix).

### Solution
Based on the proof in Part(1), we have two equations:
$$[A^{T}A]_{ij} = \sum_{k=1}^{m} [A^{T}]_{ik}[A]_{kj}$$
$$[AA^{T}]_{ij} = \sum_{k=1}^{n} [A]_{ik}[A^{T}]_{kj}$$

In order to have $A^{T}A = AA^{T}$, we need $m=n$ and $A = A^{T}$, which implies $A$ has to be a symmetric matrix.





# Problem Set 2: Matrix Factorization

Q: Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer.

###   Solution

We then can build a factorization function to decompose a square matrix into an L and U matrix.

```{r}
factorize <- function(mtx,n){
  
  # Create empty matrix for lower and upper triangular matrix 
  lower <- matrix(0, nrow=n, ncol=n)
  upper <- matrix(0, nrow=n, ncol=n)
  
  # Decomposing matrix into upper and lower triangular matrix using loop
  for (i in 1:n){
    
    # Upper triangular
    for (k in i:n){
      
      # Summation of L[i, j] * U[j, k]
      sum <- 0
      for (j in 1:i){
        sum <- sum + (lower[i,j] * upper[j,k])
      }
      
      # Evaluating U[i, k] 
      upper[i,k] <- mtx[i,k] - sum
      
    }
  
    # Lower Triangular
    for (k in i:n){
      
      if (i==k){
        lower[i,i] <- 1 # Diagonal as 1
      } else{
        
            # Summation of L[k, j] * U[j, i]
            sum <- 0
            for (j in 1:i){
              sum <- sum + (lower[k,j] * upper[j,i])
            }
            
            # Evaluating L[k, i]
            lower[k,i] = (mtx[k,i] - sum) / upper[i,i]
      }
    }
  }
  
  list <- list("U"=upper, "L"=lower)
  return(list)

}
```


#### To test the Function using a 3X3 square matrix
```{r}
# Using a sample 3x3 square matrix to test the function
mtx <- matrix(c(1,1,2,2,1,0,3,4,3), byrow=T, nrow=3)

factorize(mtx,3)
```

#### To check if LU = mtx
```{r}
# Checking if LU = mtx
factorize(mtx,3)$L %*% factorize(mtx,3)$U
```

**Hence, we say the function is valid and correct**
